Natural Stories model analyses

### 2022-03-01 Analysis/models.Rmd
Roger's in-progress version of models.Rmd

@@

```{r meaned data, eval=T, cache=T}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")

gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
gam_grnn_2 <- gam(formula_no_interact, data=grnn_mean_data, method="REML")
gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
gam_gpt_2 <- gam(formula_no_interact, data=gpt_mean_data, method="REML")

@@

write_rds(all_bootstrapped_predictions,"bootstrapped_GAM_predictions.rds")
write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")
```

```{r example bootstrap results,eval=F}
# as an example, plot effect for surprisal
summaries <- bootstrapped_gam_predictions_grnn
ggplot(summaries[[1]],aes(x=surprisal,y=rt)) + geom_line() +
  #geom_rug(aes(x=grnn_mean_data$surprisal),sides='b') +
  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_lower),linetype="dashed") +
  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_upper),linetype="dashed") +
  ylim(c(-200,600)) +
  xlab("surprisal") +
  ylab("RT contribution") +
  theme(aspect.ratio=1)
ggsave("/tmp/bootstrap.pdf")
```

>>>>

formula3_interact <- rt ~ ti(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     ti(freq,len, bs="cr")+
                     ti(surprisal,len, bs="cr")+
                     ti(surprisal,freq, bs="cr")+
                     ti(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr") +
                     ti(prev_surp, prev_len, bs="cr") +
                     ti(prev_surp, prev_freq, bs="cr")

formula2_interact <- rt ~ 
                     ti(freq,len, bs="cr")+
                     ti(surprisal,len, bs="cr")+
                     ti(surprisal,freq, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr") +
                     ti(prev_surp, prev_len, bs="cr") +
                     ti(prev_surp, prev_freq, bs="cr")
```

@@

```{r meaned data, eval=T, cache=T}
gam_ngram_1 <- gam(formula_interact, data=ngram_mean_data, method="REML")
gam_grnn_1 <- gam(formula_interact, data=grnn_mean_data, method="REML")
gam_txl_1 <- gam(formula_interact, data=txl_mean_data, method="REML")
gam_gpt_1 <- gam(formula_interact, data=gpt_mean_data, method="REML")
gam_gpt_1a <- gam(formula_interact, data=filter(gpt_mean_data,len<9), method="REML")
gam_gpt_2 <- gam(formula2_interact, data=gpt_mean_data, method="REML")
gam_gpt_3 <- gam(formula3_interact, data=gpt_mean_data, method="REML")
print(plot(getViz(gam_gpt_1x)))
print(plot(getViz(gam_gpt_3)))

gam_ngram_2 <- gam(formula_no_interact, data=ngram_mean_data, method="REML")
gam_grnn_2 <- gam(formula_no_interact, data=grnn_mean_data, method="REML")
gam_txl_2 <- gam(formula_no_interact, data=txl_mean_data, method="REML")
gam_gpt_no_interact <- gam(formula_no_interact, data=gpt_mean_data, method="REML")

@@

write_rds(all_bootstrapped_predictions,"bootstrapped_GAM_predictions.rds")
write_rds(all_surprisal_predictions,"bootstrapped_GAM_surprisal_predictions.rds")

dats_for_predictionX <- list(
  'ti(surprisal)'=dat_for_surprisal_prediction,
  'ti(freq)'=dat_for_freq_prediction,
  'ti(len)'=dat_for_length_prediction,
  'ti(prev_surp)'=dat_for_prev_surp_prediction,
  'ti(prev_freq)'=dat_for_prev_freq_prediction,
  'ti(prev_len)'=dat_for_prev_len_prediction)
system.time(res_gptX <- gam_boot(gpt_mean_data,formula2_interact,dats_for_predictionX,N=N_replicates))
bootstrapped_gam_predictions_gptX <- summarize_gam_boot_results(res_gpt,dats_for_predictionX,term_names,"GPT-2")
write_rds(res_gptX,"res_gptX.rds")
write_rds(bootstrapped_gam_predictions_gptX,"bootstrapped_gam_predictions_gptX.rds")
```

```{r example bootstrap results,eval=F}
# as an example, plot effect for surprisal
summaries <- bootstrapped_gam_predictions_gptX
summaries <- bootstrapped_gam_predictions_grnn
ggplot(summaries[[1]],aes(x=surprisal,y=rt)) + geom_line() +
  #geom_rug(aes(x=grnn_mean_data$surprisal),sides='b') +
  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_lower),linetype="dashed") +
  geom_line(aes(x=summaries[[1]]$surprisal,y=summaries[[1]]$CI_upper),linetype="dashed") +
  ylim(c(-200,600)) +
  xlab("surprisal") +
  ylab("RT contribution") +
  theme(aspect.ratio=1)
ggsave("/tmp/bootstrap.pdf")
```
